---
title: "csbewma: An R Package for the Cumulative Standardized Binomial EWMA Control Chart for Multiple Stream Processes"
abstract: >
  Monitoring binomial proportions across multiple independent streams arises routinely in
  statistical process control, with applications in manufacturing, healthcare, cybersecurity,
  finance, and environmental monitoring. Exponentially weighted moving average (EWMA) charts are
  attractive in this setting for their sensitivity to small and moderate shifts, but existing
  EWMA implementations for binomial data rely on asymptotic variance approximations that are
  inaccurate during the early phase of monitoring, when the sample history is short. This paper
  introduces csbewma, an R package implementing the Cumulative Standardized Binomial EWMA
  (CSB-EWMA) control chart, which uses an exact, closed-form, time-varying variance for the EWMA
  statistic under a binomial data-generating process. This enables adaptive control limits that
  are valid from the first observation rather than only asymptotically. Beyond the charting
  engine, csbewma provides a post-hoc identification framework for determining which stream or
  streams triggered a signal, using exact binomial p-values with Bonferroni, Holm, or
  Benjamini-Hochberg multiplicity correction, along with a set of ggplot2-based plotting
  functions for control charts and diagnostic displays. The package's functions and S3 class
  design are described, its usage is illustrated with real console output and control charts,
  and the simulation evidence behind its default settings is summarized. csbewma is distributed
  under the MIT license and is available from the Comprehensive R Archive Network.
draft: yes
author:
  - name: Faruk Muritala
    affiliation: Kennesaw State University
    address:
    - School of Data Science and Analytics, College of Computing and Software Engineering
    - Kennesaw, Georgia, United States of America
    email: fmurital@students.kennesaw.edu
    orcid: 0000-0002-5857-9874
  - name: Austin Brown
    affiliation: Kennesaw State University
    address:
    - School of Data Science and Analytics, College of Computing and Software Engineering
    - Kennesaw, Georgia, United States of America
    orcid: 0000-0003-1530-0009
  - name: Dhrubajyoti Ghosh
    affiliation: Kennesaw State University
    address:
    - School of Data Science and Analytics, College of Computing and Software Engineering
    - Kennesaw, Georgia, United States of America
    orcid: 0000-0002-3360-3786
  - name: Sherry Ni
    affiliation: Kennesaw State University
    address:
    - School of Data Science and Analytics, College of Computing and Software Engineering
    - Kennesaw, Georgia, United States of America
    orcid: 0000-0003-0634-0025
type: package
output:
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: RJreferences.bib
---

```{r setup, include=FALSE}
# Every eval=TRUE chunk in this document runs for real at knit time -- nothing in the
# "Illustrative use" section is hand-typed output. Every console message, print() block, and
# figure there is produced live by knitr, per the R Journal's requirement that R outputs come
# from an evaluation of R code rather than manual transcription.
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.align = "center"
)
```

```{r load-packages, include=FALSE}
# kableExtra is a build-time dependency of this .Rmd (used to fit wide tables to the page width
# below); it is not a dependency of the csbewma package itself.
library(csbewma)
library(ggplot2)
library(kableExtra)
library(patchwork)
```

```{r figure-label-patch, include=FALSE}
# Local, display-only override for this article's build. The CRAN release of csbewma (v1.0.1)
# draws the "Signal at t = ..." annotation in plot.csb_ewma() and plot_csb_ewma_direct() with
# hjust = 0, which left-aligns the label starting at the signal point; because the signal in this
# article's examples occurs near the right end of the plotted series, the label can run past the
# panel's right edge. Right-aligning it (hjust = 1) keeps the label inside the panel. This changes
# label justification only: no data, statistic, axis, or numerical result shown anywhere in this
# article is affected. It is filed as a fix for the package's next CRAN release so that
# plot.csb_ewma() and plot_csb_ewma_direct() draw the label this way by default going forward; see
# README-FOR-AUTHOR.md for the corresponding upstream change to plot.R.
plot.csb_ewma <- function(x, title = "CSB-EWMA Control Chart", show_signal = TRUE, ...) {
  library(ggplot2)
  if (!inherits(x, "csb_ewma")) stop("Object must be of class 'csb_ewma'")
  t_values <- 1:x$T_sig
  plot_data <- data.frame(t = t_values, r = x$r_history, UCL = x$UCL_history, LCL = x$LCL_history)
  p <- ggplot(plot_data, aes(x = t)) +
    geom_line(aes(y = r, color = "EWMA Statistic"), linewidth = 1) +
    geom_line(aes(y = UCL, color = "UCL"), linetype = "dashed", linewidth = 0.8) +
    geom_line(aes(y = LCL, color = "LCL"), linetype = "dashed", linewidth = 0.8) +
    geom_hline(yintercept = 0, linetype = "dotted", color = "gray50", linewidth = 0.5) +
    labs(title = title, x = "Time Point (t)", y = expression(r[t] ~ "(EWMA Statistic)"), color = "Series") +
    scale_color_manual(values = c("EWMA Statistic" = "#1F7AB5", "UCL" = "#D62828", "LCL" = "#D62828")) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom", legend.title = element_text(face = "bold", size = 10),
          legend.text = element_text(size = 9), plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
          axis.title = element_text(face = "bold", size = 11), axis.text = element_text(size = 9),
          panel.grid.minor = element_blank(), panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
          panel.background = element_rect(fill = "white", color = NA), plot.background = element_rect(fill = "white", color = NA))
  if (show_signal && x$signal_detected) {
    signal_data <- data.frame(t = x$signal_time, r = x$r_history[x$signal_time])
    p <- p + geom_point(data = signal_data, aes(x = t, y = r), color = "#D62828", size = 4.5, shape = 18) +
      annotate("text", x = x$signal_time, y = x$r_history[x$signal_time] + 0.15,
               label = paste("Signal at t =", x$signal_time), color = "#D62828", size = 3.5,
               fontface = "bold", hjust = 1)
  }
  print(p)
  invisible(p)
}

plot_csb_ewma_direct <- function(result, title = "CSB-EWMA Control Chart", show_signal = TRUE) {
  library(ggplot2)
  t_values <- 1:result$T_sig
  plot_data <- data.frame(t = t_values, r = result$r_history, UCL = result$UCL_history, LCL = result$LCL_history)
  p <- ggplot(plot_data, aes(x = t)) +
    geom_line(aes(y = r, color = "EWMA Statistic"), linewidth = 1) +
    geom_line(aes(y = UCL, color = "UCL"), linetype = "dashed", linewidth = 0.8) +
    geom_line(aes(y = LCL, color = "LCL"), linetype = "dashed", linewidth = 0.8) +
    geom_hline(yintercept = 0, linetype = "dotted", color = "gray50", linewidth = 0.5) +
    labs(title = title, x = "Time Point (t)", y = expression(r[t] ~ "(EWMA Statistic)"), color = "Series") +
    scale_color_manual(values = c("EWMA Statistic" = "#1F7AB5", "UCL" = "#D62828", "LCL" = "#D62828")) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom", legend.title = element_text(face = "bold", size = 10),
          legend.text = element_text(size = 9), plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
          axis.title = element_text(face = "bold", size = 11), axis.text = element_text(size = 9),
          panel.grid.minor = element_blank(), panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
          panel.background = element_rect(fill = "white", color = NA), plot.background = element_rect(fill = "white", color = NA))
  if (show_signal && result$signal_detected) {
    p <- p + geom_point(data = data.frame(t = result$signal_time, r = result$r_history[result$signal_time]),
                        aes(x = t, y = r), color = "#D62828", size = 4.5, shape = 18) +
      annotate("text", x = result$signal_time, y = result$r_history[result$signal_time] + 0.15,
               label = paste("Signal at t =", result$signal_time), color = "#D62828", size = 3.5,
               fontface = "bold", hjust = 1)
  }
  print(p)
  invisible(p)
}

plot_chart_with_flagged <- function(chart_result, flagged_results, layout = "side") {
  library(ggplot2)
  library(patchwork)
  p1 <- plot_csb_ewma_direct(chart_result, show_signal = TRUE)
  p2 <- plot_flagged_streams(flagged_results, alpha = 0.05)
  if (layout == "side") {
    combined_plot <- p1 + p2 + plot_layout(ncol = 2, widths = c(1, 0.8))
  } else {
    combined_plot <- p1 / p2 + plot_layout(nrow = 2, heights = c(1, 0.8))
  }
  combined_plot <- combined_plot +
    plot_annotation(title = "CSB-EWMA Diagnostic Dashboard",
                     theme = theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 16)))
  print(combined_plot)
  invisible(combined_plot)
}
```

## Introduction

Statistical process control is a foundational methodology for monitoring, controlling, and
improving processes across manufacturing, healthcare, finance, and service industries
[@montgomery2019]. Since Shewhart introduced the control chart in the 1920s [@shewhart1931], the
central task of statistical process control has been to separate common cause variation, which is
inherent to a stable process, from special cause variation that signals a change requiring
intervention. Traditional Shewhart charts are effective for detecting large shifts but are
comparatively insensitive to small and moderate process changes. This motivated memory-type
alternatives such as the cumulative sum (CUSUM) chart [@page1954] and the exponentially weighted
moving average (EWMA) chart [@roberts1959], which incorporate the full process history through
exponential weighting and are prized for their sensitivity to small, persistent shifts
[@lucas1990].

A growing class of applications requires monitoring several parallel data streams at once rather
than a single process, so-called multiple stream processes, which originated in manufacturing
settings such as multi-head filling machines and multi-spindle operations [@boyd1950]. A more
recent line of work pursues distribution-free multiple stream process monitoring by dichotomizing
each stream's observations relative to an in-control median, reducing the monitoring problem to
tracking a binomial proportion. The nonparametric extended median test CUSUM chart of
[@brown2021] is the primary contribution of this kind, though it focuses on detecting
comparatively large shifts and relies on asymptotic properties that are less suitable for
small-sample, early-phase monitoring.

Applying the more sensitive EWMA framework to binomial multiple stream data exposes a genuine
theoretical gap. Existing EWMA charts for binomial or count data [@gan1990; @gan1993;
@anastasopoulou2022] rely on an asymptotic variance approximation that does not capture the
time-varying nature of the true variance during the initial monitoring phase, when data are
limited. Because most implementations then use constant control limits derived from this
asymptotic variance, they either signal falsely too often or fail to exploit the increasing
precision that accrues with each new observation.

This gap is closed by deriving the exact, closed-form, time-varying mean and variance of the EWMA
statistic for cumulative binomial counts observed across $k$ independent streams [@muritala2026a;
@muritala2026b]. The resulting Cumulative Standardized Binomial EWMA (CSB-EWMA) chart uses these
exact moments to construct adaptive control limits that are statistically valid from the very
first observation, rather than only in the large-sample limit, while retaining the
distribution-free property of the underlying dichotomization approach.

This paper documents the csbewma package, the R implementation of that methodology, with an
emphasis on how to use it. The Statistical methodology section summarizes the theory so that the
package's arguments and defaults make sense. The csbewma package section describes the package's
design and functions. The Illustrative use section is the core usage illustration: real code and
the real console output and control charts it produces when this document is rendered. The Chart
parameter selection section summarizes the simulation evidence behind the recommended $\lambda$
and $L$ settings. The Post-hoc identification section documents the post-hoc identification
methodology, its rationale, its simulated performance, and a worked illustration of its
early-signal limits. The Limitations section discusses the method's limitations, and the final
section concludes.

## Statistical methodology

This section summarizes, at a level sufficient to understand the package's arguments, the
statistical derivation underlying the CSB-EWMA chart. Full derivations and proofs appear in
[@muritala2026a] and Chapter 3 of [@muritala2026diss].

### Process framework

The chart monitors $k$ mutually independent data streams. Within a given stream, observations
collected at successive sampling times are also assumed mutually independent, and the process
must be measurable on at least an ordinal scale [@brown2021]. Let $y_{ij}$ denote the $j$th
observation from stream $i$, $i = 1, \ldots, k$, which has an unknown probability distribution but
a known, stable in-control median. A binary indicator $x_{ij}$ is defined as $1$ if $y_{ij}$
exceeds that median and $0$ otherwise. Provided stream $i$'s median remains stable, $x_{ij}$
follows a $\mathrm{Binomial}(n=1, p_0=0.5)$ distribution. Because the streams, and within a time
point the observations across streams, are independent, the column total $C_j$ (the sum of
$x_{ij}$ over streams $i=1,\ldots,k$) follows $\mathrm{Binomial}(n=k, p_0=0.5)$, and the cumulative
count of exceedances through time $t$, $Q_t$ (the running sum of $C_j$), follows
$\mathrm{Binomial}(n=kt, p_0=0.5)$. The standardized statistic is:

$$
Q_t \sim \mathrm{Binomial}(kt,\, p_0), \qquad
W_t = \frac{Q_t - tk p_0}{\sqrt{tk\,p_0(1-p_0)}}
$$

with mean $0$ and variance $1$ by construction.

### EWMA recursion and its exact moments

The EWMA statistic follows the recursion below, with $r_0 = 0$ in the package and smoothing
parameter $0 < \lambda \le 1$:

$$
r_t = \lambda W_t + (1-\lambda) r_{t-1}, \qquad r_0 = 0, \qquad 0 < \lambda \le 1
$$

[@muritala2026a] show that $E[r_t] = (1-\lambda)^t r_0$, which is exactly zero in the package's
implementation since $r_0 = 0$. The exact variance follows from the covariance structure
$\mathrm{Cov}(W_i, W_j) = \sqrt{\min(i,j)}/\sqrt{\max(i,j)}$ and is given, for finite $t$, by:

$$
\operatorname{Var}(r_t) \;=\; \lambda^2 \left[\, 2\sum_{j=1}^{t-1}\sum_{i=j+1}^{t} (1-\lambda)^{\,2t-i-j}\sqrt{\tfrac{j}{i}} \;+\; \sum_{i=1}^{t} (1-\lambda)^{\,2t-2i} \,\right]
$$

This is precisely what `var_rt_exact_single()` computes, as described in the Exact variance
computation section below.

### Asymptotic behavior and adaptive control limits

As $t$ grows large, $\operatorname{Var}(r_t)$ approaches $1$ [@muritala2026a, Theorem 3]. The
package therefore uses time-varying control limits, where $L$ is the control limit multiplier:

$$
UCL_t = L\sqrt{\operatorname{Var}(r_t)}, \qquad LCL_t = -L\sqrt{\operatorname{Var}(r_t)}
$$

These limits are narrow near $t=1$ and widen toward $\pm L$ as $t$ grows, which is the mechanism
by which the chart avoids both the elevated early-phase false alarm risk and the loss of
early-phase sensitivity that occur when a chart instead uses constant, asymptotic limits from the
outset.

### Monte Carlo validation

[@muritala2026a] and Chapter 3 of [@muritala2026diss] validate the closed-form expressions against
10,000 Monte Carlo replications under the in-control process, with $k=10$ streams, $p_0=0.5$, and
$\lambda=0.2$ (R version 4.3.2; @rcore2023). Table 1 reproduces the reported
validation metrics; these are cited results from that Monte Carlo study, not values recomputed by
this document (a 10,000-replication study is outside the R Journal's ten-minute knit budget), so
the table below is built from a small data frame of those reported numbers rather than from a live
simulation.

```{r validation-table, echo=FALSE}
validation_df <- data.frame(
  Time = c(10, 50, 100, 500, 1000),
  `Theoretical Mean` = c(0.000, 0.000, 0.000, 0.000, 0.000),
  `Simulated Mean` = c(-5.54e-3, -5.23e-3, -1.95e-3, -6.06e-4, -3.09e-3),
  `Theoretical Var.` = c(0.6369, 0.9512, 0.9768, 0.9955, 0.9978),
  `Simulated Var.` = c(0.6385, 0.9359, 0.9758, 0.9967, 0.9996),
  `Rel. Bias (Var.)` = c("0.26%", "-1.61%", "-0.10%", "0.12%", "0.19%"),
  check.names = FALSE
)
knitr::kable(
  validation_df,
  caption = "Validation of the exact EWMA mean and variance against Monte Carlo simulation (k = 10, p0 = 0.5, lambda = 0.2, 10,000 replications). Source: Muritala et al. (2026a, Table 4.1); Muritala (2026, Table 3.2).",
  booktabs = TRUE
)
```

The simulated means are effectively zero throughout, with a reported root mean square bias of
$2.89\times10^{-3}$. The exact variance formula tracks the simulated variance closely at every
reported time point, with relative bias below 2% throughout, converging to about 0.19% by
$t=1000$. The variance reaches 99% of its asymptotic value by $t=227$, which is the basis for the
package's default `converge_t = 500` cutoff described below.

## The csbewma package

The csbewma package is distributed on CRAN as version 1.0.1 (published 2026-06-05; DOI
[10.32614/CRAN.package.csbewma](https://doi.org/10.32614/CRAN.package.csbewma)). Per its
DESCRIPTION file, it requires R (>= 3.5), imports ggplot2, stats, and patchwork, suggests testthat
(>= 3.0.0), and is distributed under the MIT license. This paper documents eleven exported
functions, plus the `print` and `plot` S3 methods for class `csb_ewma`, grouped by role in Table
2.

```{r functions-table, echo=FALSE}
functions_df <- data.frame(
  Group = c(
    "Exact variance / control limits",
    "Charting engine",
    "Post-hoc identification",
    "Visualization"
  ),
  `Function(s)` = c(
    "var_rt_exact_single(), precompute_variance()",
    "run_csb_ewma(), csb_ewma(), print.csb_ewma()",
    "calculate_pvalues(), apply_multiple_testing(), identify_ooc(), flagged_streams_summary()",
    "plot.csb_ewma(), plot_csb_ewma_direct(), plot_flagged_streams(), plot_chart_with_flagged()"
  ),
  Purpose = c(
    "Compute Var(r_t) exactly at a single t, or precompute a full variance vector with asymptotic fallback beyond converge_t.",
    "Low-level chart algorithm on a binary matrix; user-facing wrapper handling variance precomputation and post-hoc identification; S3 print method.",
    "Exact binomial p-values per stream; Bonferroni, Holm, BH, or raw multiplicity correction via p.adjust(); combined stream-level identification; summary of flagged streams.",
    "ggplot2 control chart with signal annotation; direct, non-S3 equivalent; bar chart of stream p-values; patchwork dashboard combining the chart and the bar chart."
  ),
  check.names = FALSE
)
knitr::kable(
  functions_df,
  caption = "Exported functions and S3 methods of csbewma 1.0.1, by functional group.",
  booktabs = TRUE
) %>%
  kableExtra::column_spec(1, width = "1in") %>%
  kableExtra::column_spec(2, width = "1.8in") %>%
  kableExtra::column_spec(3, width = "2.6in") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

The Illustrative use section below demonstrates the resulting workflow with real package output.

### Exact variance computation

`var_rt_exact_single(lambda, t)` evaluates the closed-form expression above for a single time
point, returning $0$ when $t=0$. The function accumulates two pieces exactly as in the derivation:
an off-diagonal contribution, summed for $j=1,\ldots,t-1$ using vectorized weights
$(1-\lambda)^{2t-j-i}$ for $i$ running from $j+1$ to $t$ and the covariance terms $\sqrt{j/i}$,
doubled for the symmetric triangle, and a diagonal contribution $(1-\lambda)^{2t-2i}$ summed over
$i=1,\ldots,t$. The total is multiplied by $\lambda^2$. Because this double summation costs
$O(t^2)$ if evaluated for every $t$ up to a long monitoring horizon, `precompute_variance(lambda,
max_t, converge_t = 500)` instead computes the exact variance only for $t = 1, \ldots,
\min(\texttt{converge\_t}, \texttt{max\_t})$ and fills the remaining entries, for $t$ greater than
`converge_t`, with the asymptotic value 1. The default `converge_t = 500` is set well above the
empirical 99% convergence point of $t=227$ established above; the function's own documentation
describes it as safe and efficient. The resulting numeric vector, one entry per time point, is the
`var_cache` consumed by `run_csb_ewma()`.

### The charting functions

`run_csb_ewma(bin_matrix, lambda, L, var_cache, max_time = NULL, p0 = 0.5)` is the core algorithm,
implemented, in the words of its own documentation, exactly as in the original simulation code.
After confirming that `bin_matrix` contains only 0s and 1s, that $0 < \lambda \le 1$, and that
$L>0$, it sets $k = \texttt{nrow(bin\_matrix)}$, $\mu_0 = kp_0$, and $\sigma^2_0 = kp_0(1-p_0)$. It
then steps through $t=1,\ldots,\texttt{max\_time}$, at each step summing the binary indicators
across streams for that time point ($C_t$), accumulating the running total (`cum_sum`), forming
the standardized statistic $W_t = (\texttt{cum\_sum} - \mu_0 t)/\sqrt{t\sigma^2_0}$, updating the
EWMA $r_t$, and comparing $r_t$ against the exact control limits above. The loop stops and records
`signal_time` the first time $r_t$ exceeds either limit. If no signal occurs by `max_time`, the
function issues a warning and sets `T_sig` to `max_time`. It returns an object of S3 class
`csb_ewma` holding `signal_time`, `signal_detected`, `T_sig`, the `r_history`, `UCL_history`, and
`LCL_history` vectors truncated to `T_sig`, the truncated `bin_matrix`, per-stream successes, $k$,
$\lambda$, $L$, and $p_0$.

`csb_ewma(data, lambda, L, p0 = 0.5, max_time = NULL, posthoc_method = "BH", alpha = 0.05)` is the
function most users call directly. It validates that `data` is a matrix with $0 < \lambda \le 1$
and $L>0$, precomputes the variance vector with `precompute_variance()`, and runs
`run_csb_ewma()`. If a signal is detected, it automatically performs post-hoc identification by
calling `identify_ooc()` on the truncated binary matrix with the requested `posthoc_method`
(Benjamini-Hochberg, `"BH"`, by default) and `alpha`, attaching the resulting data frame as
`$flagged` along with `$posthoc_method` and `$alpha`. It also prints short status messages as it
runs, for example confirming that the input is binary, reporting the signal time, and reporting
whether any streams were flagged, which the Illustrative use section reproduces live. `print.csb_ewma()`
then displays $\lambda$, $L$, $p_0$, and $k$, whether and when a signal occurred, and, when
post-hoc results are attached, the post-hoc method, alpha, the number of flagged streams, and
their stream numbers if any were flagged.

`csb_ewma()` takes `data` as an already binary $0/1$ matrix; it does not accept a `distribution`
argument, as a direct console session confirms. The `data` argument is `bin_matrix`, a $k$ by $t$
matrix with one row per stream and one column per sampling time, entries $x_{ij}\in\{0,1\}$,
formed exactly as in the process framework described above: $x_{ij}=1$ if the $j$th observation
from stream $i$ exceeds that stream's known, stable in-control median (or otherwise indicates the
monitored event), and $0$ otherwise. Table 3 lays out this structure
explicitly.

```{r dataformat-table, echo=FALSE}
dataformat_df <- data.frame(
  Stream = c("1", "2", "...", "k", "Column totals"),
  `j = 1` = c("x_11", "x_21", "...", "x_k1", "C_1"),
  `j = 2` = c("x_12", "x_22", "...", "x_k2", "C_2"),
  `...` = c("...", "...", "...", "...", "..."),
  `j = t` = c("x_1t", "x_2t", "...", "x_kt", "C_t"),
  check.names = FALSE
)
knitr::kable(
  dataformat_df,
  caption = "Required input structure for csb\\_ewma(): a k stream by t sample recoded binary matrix, with column totals $C_j$.",
  booktabs = TRUE,
  align = "lcccc"
)
```

Because it is assumed that, for a given sampling time $j$, the observations across the $k$ streams
are mutually independent Bernoulli random variables, the column total $C_j = \sum_i x_{ij}$
follows $\mathrm{Binomial}(n=k, p_0=0.5)$ under the in-control process, exactly as introduced
above; `csb_ewma()` forms $C_j$ internally for every column of `bin_matrix` and accumulates it
into the running total $Q_t$. Continuous or otherwise non-binary measurements must be recoded by
the user into this form, by whatever thresholding rule suits the application, before calling
`csb_ewma()`.

### Post-hoc identification of out-of-control streams

Detecting that the aggregate chart has signaled does not by itself indicate which of the $k$
streams is responsible. csbewma addresses this with a post-hoc module that its own source file
describes as taken directly from the original simulation code underlying [@muritala2026diss].
`calculate_pvalues(bin_matrix, p0 = 0.5)` computes, for every stream $i$ with $S_i$ successes out
of $T_{sig} = \texttt{ncol(bin\_matrix)}$ observations, the exact upper-tail binomial p-value $p_i
= 1 - \texttt{pbinom}(S_i - 1, \texttt{size} = T_{sig}, \texttt{prob} = p_0)$, that is,
$P(\mathrm{Binomial}(T_{sig}, p_0) \ge S_i)$. `apply_multiple_testing(pvals, method = "BH", alpha
= 0.05)` adjusts these raw p-values with R's built-in `p.adjust()`, supporting `"bonferroni"`,
`"holm"`, `"BH"`, or `"raw"` (no adjustment). `identify_ooc(bin_matrix, alpha = 0.05, method =
"BH", p0 = 0.5)` wraps both steps and returns a data frame with columns `stream`, `successes`,
`proportion`, `p_value`, `adjusted_p_value`, and `flagged`, sorted with flagged streams first and
then by increasing p-value. `flagged_streams_summary(flagged_results)` extracts just the flagged
rows and their count.

The default `posthoc_method = "BH"` reflects a specific, simulation-based recommendation rather
than an arbitrary choice, which the Post-hoc identification section below summarizes.

### Visualization

`plot.csb_ewma(x, title = "CSB-EWMA Control Chart", show_signal = TRUE, ...)` is the S3 plot
method dispatched by `plot()` on a `csb_ewma` object; `plot_csb_ewma_direct()` draws the identical
chart without S3 dispatch. Both plot the EWMA trajectory $r_t$ as a solid blue line, the
time-varying UCL and LCL as dashed red lines, a dotted gray zero line, and, when `show_signal =
TRUE` and a signal was detected, a red diamond marker with a "Signal at $t = \ldots$" annotation,
using a clean `theme_minimal()` ggplot2 theme with a bottom legend. `plot_flagged_streams(flagged_results,
alpha = 0.05, title = "Stream P-values")` draws a bar chart of the negative log base ten p-value
for each stream, coloring flagged streams red and unflagged streams gray, with a dashed reference
line at the corresponding threshold. `plot_chart_with_flagged(chart_result, flagged_results,
layout = "side")` combines the control chart and the flagged stream bar chart into one patchwork
figure, either side by side or stacked, under a common title.

## Illustrative use of the package

This section shows the package running end to end. Unlike a typical journal article, nothing
below is a transcript pasted in by hand: every code chunk in this section is evaluated live when
this document is rendered, so the console messages, the `print()` output, and the figures are
produced by your own installation of csbewma at knit time. Because every random draw below is
seeded, re-knitting this document should reproduce the same signal times and flagged streams
described in the surrounding prose.

### Running the chart on in-control data

```{r example1-run}
set.seed(123)
bin_data <- matrix(rbinom(10 * 100, 1, 0.5), nrow = 10, ncol = 100)
result <- csb_ewma(bin_data, lambda = 0.175, L = 1.375)
```

`bin_data` is a 10-stream by 100-time-point in-control binary matrix, each entry an independent
Bernoulli(0.5) draw. Calling `csb_ewma()` with $(\lambda, L) = (0.175, 1.375)$, the smallest
$\lambda$ pair in Table 4 below, printed the status messages shown above as it
ran. The chart signaled even though the data were generated in-control throughout, which is
expected: with $L=1.375$ the chart targets an in-control average run length of about 370 (Table
4), so an early in-control false alarm is a plausible draw from that distribution
rather than a sign of a problem. Because a signal was detected, `csb_ewma()` automatically ran the
Benjamini-Hochberg post-hoc procedure.

### Inspecting the result

```{r example1-print}
print(result)
```

This is exactly the format implemented by `print.csb_ewma()`: chart parameters, the signal
detection outcome, and, because `$flagged` is attached, the post-hoc method, alpha, and flagged
stream count.

### Plotting the chart

```{r example1-plot, fig.cap="Output of plot(result) for the in-control example above, produced by the package's own plot.csb\\_ewma() method."}
plot(result)
```

The solid blue line is the EWMA trajectory $r_t$; the dashed red lines are the time-varying UCL
and LCL described above, visibly narrow near $t=1$ and widening toward their asymptotic values as
$t$ grows; the red diamond marks the signal.

### A second example: detecting and identifying a real shift

The first example was fully in control. This second example introduces genuine shifts in three of
ten streams and shows the chart detecting the shift and the post-hoc procedure identifying which
streams caused it.

```{r example2-run}
set.seed(2026)
k <- 10          # Number of binary streams/processes
n_time <- 500    # Number of observations over time
p0 <- 0.5        # In-control success probability
bin_matrix <- matrix(rbinom(k * n_time, size = 1, prob = p0), nrow = k, ncol = n_time)

# Shift the first three streams to elevated probabilities
p1_vector <- c(0.56, 0.55, 0.60, rep(0.5, 7))
for (i in 1:3) {
  bin_matrix[i, ] <- rbinom(n_time, size = 1, prob = p1_vector[i])
}

var_cache <- precompute_variance(lambda = 0.175, max_t = n_time)

result2 <- csb_ewma(
  data = bin_matrix,
  lambda = 0.175,
  L = 1.375,
  p0 = p0,
  posthoc_method = "BH",
  alpha = 0.05
)
```

Streams 1 through 3 have their success probability raised from $p_0=0.5$ to $0.56$, $0.55$, and
$0.60$ respectively, a small to moderate shift by the $\delta$ scale used in the Chart parameter
selection section below; the remaining seven streams stay in control. Precomputing the variance
vector explicitly with `precompute_variance()` before calling `csb_ewma()` is optional, since
`csb_ewma()` does this internally, but is shown here for illustration.

```{r example2-print}
print(result2)
```

```{r example2-plot, fig.cap="Output of plot(result2) for the shifted-stream example. The chart signals where the EWMA trajectory crosses the upper control limit."}
plot(result2)
```

Calling `identify_ooc()` directly on the full 500-column matrix, rather than only on the data up
to the signal time that `csb_ewma()` uses automatically, illustrates the full data frame the
post-hoc functions return:

```{r example2-identify}
full_results <- identify_ooc(bin_matrix, p0 = p0, alpha = 0.05, method = "BH")
knitr::kable(
  full_results,
  caption = "identify\\_ooc(bin\\_matrix, p0 = 0.5, alpha = 0.05, method = \"BH\") applied to the full 500-column matrix constructed above, sorted with flagged streams first.",
  booktabs = TRUE,
  row.names = FALSE
)
```

With the full 500-column history available, the genuinely shifted streams are flagged, including
any that the automatic post-hoc step at the earlier signal time had not yet flagged. This is a
direct illustration of the point developed formally in the Illustration subsection below:
attribution power depends on how much post-signal data is available, and improves as more data
accumulates, which is exactly why `csb_ewma()`'s built-in post-hoc step and a later, separate call
to `identify_ooc()` on more data can legitimately disagree.

### Visualizing flagged streams

Two visualization functions from Table 2 and the Visualization subsection above, `plot_flagged_streams()`
and `plot_chart_with_flagged()`, have not yet been shown running. Both take the data frame returned
by `identify_ooc()` as their `flagged_results` argument. `plot_flagged_streams(flagged_results,
alpha = 0.05, title = "Stream P-values")` draws the per-stream $-\log_{10}(p)$ bar chart described
above; applied to `full_results` from the full 500-column history just computed, it shows all
three genuinely shifted streams in red:

```{r example2-flagged-plot, fig.cap="Output of plot\\_flagged\\_streams(full\\_results, alpha = 0.05) for the second example above, using the full 500-column identify\\_ooc() results. Streams 1, 2, and 3, the genuinely shifted streams, are shown in red; unflagged streams are gray; the dashed line marks the alpha = 0.05 reference threshold on the -log10(p) scale."}
plot_flagged_streams(full_results, alpha = 0.05)
```

`plot_chart_with_flagged(chart_result, flagged_results, layout = "side")` combines a control chart
and a flagged-stream bar chart into one patchwork figure. Internally it builds this by calling
`plot_csb_ewma_direct()` and `plot_flagged_streams()` in turn before combining their output, so a
live call displays the two component plots individually before the combined dashboard; only the
final, combined figure is kept below (`fig.keep = "last"`), since that combined dashboard is the
function's intended result. Pairing `result2`, the chart object from the signal at $t=33$, with
`full_results`, the fuller post-hoc picture from the full 500-column history, puts the original
signal-time chart side by side with the more complete attribution that only the additional
post-signal data makes possible, the same point made in prose above:

```{r example2-dashboard-plot, fig.keep="last", fig.cap="Output of plot\\_chart\\_with\\_flagged(result2, full\\_results, layout = \"side\") for the second example above, pairing the original control chart (signal at t = 33) with the flagged-stream bar chart computed from the fuller 500-column history."}
plot_chart_with_flagged(result2, full_results, layout = "side")
```

## Chart parameter selection and detection performance

The default $\lambda$ and $L$ values suggested throughout this paper, and the recommendations
below, are drawn from the simulation study reported in [@muritala2026b] and Chapter 4 of
[@muritala2026diss]. This section summarizes that study's design and findings, as cited results,
so that users can choose parameters appropriate to their own application; reproducing the full
161,040-combination factorial grid described below is well outside the ten-minute budget for
knitting this article, so, consistent with the R Journal's own guidance for computationally
intensive methods, the tables here are built from the study's reported summary numbers rather than
recomputed live.

### Simulation design

A full factorial design was used to identify $(\lambda, L)$ pairs achieving target in-control
average run lengths (ARL0) of 370, the analogue of traditional three-sigma limits, and 500, a
stricter false alarm target, following [@montgomery2019] and [@lucas1990]. The design varied the
smoothing constant $\lambda$ from 0.10 to 0.90 in steps of 0.025, the control limit multiplier $L$
from 1.00 to 2.50 in steps of 0.025, the number of streams $k$ over $\{2,3,5,10,15,20,50,100\}$,
and the shift magnitude $\delta$ from 0.05 to 0.50 in steps of 0.05, with the in-control
proportion fixed at $p_0=0.5$. This yields $33 \times 61 \times 8 \times 10 = 161{,}040$ unique
parameter combinations, each evaluated with 10,000 Monte Carlo replications for ARL0 calibration
and 50,000 replications for ARL1 estimation, computed in R version 4.3.2 [@rcore2023].

### Optimal (lambda, L) pairs

Table 4 reports representative optimal $(\lambda, L)$ combinations across the
$\lambda$ range, for both ARL0 targets. A direct, near-linear positive relationship holds between
$\lambda$ and $L$: smaller $\lambda$ values, which place more weight on the process history and
produce a smoother statistic, require a correspondingly smaller $L$ to hold the same target ARL0.

```{r optimal-table, echo=FALSE}
optimal_df <- data.frame(
  `lambda range` = c("0.1-0.2", "0.2-0.3", "0.3-0.4", "0.4-0.5", "0.5-0.6", "0.6-0.7", "0.7-0.8", "0.8-0.9", "0.9-1.0"),
  `Opt. lambda (370)` = c(0.175, 0.200, 0.350, 0.400, 0.550, 0.650, 0.725, 0.800, 0.900),
  `Opt. L (370)` = c(1.375, 1.400, 1.500, 1.525, 1.575, 1.625, 1.650, 1.675, 1.700),
  `Achieved ARL0 (370)` = c(372.32, 375.21, 373.90, 374.48, 368.83, 373.83, 372.99, 367.53, 360.11),
  `Opt. lambda (500)` = c(0.150, 0.200, 0.300, 0.450, 0.525, 0.650, 0.700, 0.850, 0.925),
  `Opt. L (500)` = c(1.525, 1.575, 1.650, 1.725, 1.750, 1.800, 1.800, 1.875, 1.875),
  `Achieved ARL0 (500)` = c(501.31, 500.80, 501.11, 500.69, 501.07, 505.32, 502.93, 501.88, 501.88),
  check.names = FALSE
)
knitr::kable(
  optimal_df,
  caption = "Optimal EWMA chart design parameters by smoothing constant range, for target ARL0 = 370 and ARL0 = 500. Source: Muritala et al. (2026b, Table 2); Muritala (2026, Table 4.3.2).",
  booktabs = TRUE
) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

### Out-of-control detection and robustness across distributions

Detection speed was evaluated as ARL1, the expected number of samples to signal under a sustained
shift of magnitude $\delta$ from $p_0=0.5$ to $p_0+\delta$, using continuous data from four
distributions, Normal, Laplace, Uniform, and Exponential, each dichotomized at its in-control
median before charting. For small shifts ($\delta=0.05$), detection required 23 to 30 samples on
average for the ARL0 = 370 target and 30 to 36 samples for the ARL0 = 500 target. At the moderate
shift $\delta=0.20$, ARL1 dropped to 3 to 6 samples, and for large shifts ($\delta \ge 0.30$) most
combinations achieved ARL1 of 3 or fewer, several reaching the theoretical minimum of 1.

```{r arl1-table, echo=FALSE}
arl1_df <- data.frame(
  `(lambda, L)` = c("0.175, 1.375", "0.200, 1.400", "0.350, 1.500", "0.400, 1.525", "0.550, 1.575", "0.650, 1.625", "0.725, 1.650", "0.800, 1.675", "0.900, 1.700"),
  Normal = c(6, 5, 4, 4, 3, 3, 3, 3, 3),
  Laplace = c(6, 5, 4, 4, 3, 3, 3, 3, 3),
  Uniform = c(6, 5, 4, 4, 3, 3, 3, 3, 3),
  Exponential = c(6, 5, 4, 4, 3, 3, 3, 3, 3),
  `Max. difference` = c(0, 0, 0, 0, 0, 0, 0, 0, 0),
  check.names = FALSE
)
knitr::kable(
  arl1_df,
  caption = "ARL1 at delta = 0.20 (target ARL0 = 370) across four data-generating distributions, for each optimal (lambda, L) pair above. Source: Muritala et al. (2026b, Sec. 3.4); Muritala (2026, Table 4.3.3).",
  booktabs = TRUE
) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

The maximum difference in ARL1 across the four distributions is zero for every reported
combination, and the coefficient of variation of ARL1 across distributions stays below 0.10
across the range of shift magnitudes examined (0.065 to 0.090 at the smallest detectable shift,
$\delta=0.05$), a level conventionally read as excellent consistency in statistical quality
control. Together these results indicate that detection performance depends on the dichotomized
exceedance rate rather than on the shape of the underlying continuous distribution, supporting the
chart's distribution-free design.

### Practical recommendations

Parameter selection should be guided by the expected shift magnitude and the desired false alarm
rate. To detect small to moderate shifts ($\delta < 0.20$), choose $\lambda$ in the range 0.1 to
0.3; for larger shifts ($\delta \ge 0.30$), values between 0.3 and 0.6 balance sensitivity and
robustness. When the shift size is unknown, $\lambda = 0.2$ is a reasonable default. $L$ should be
chosen inversely to $\lambda$ to hold the target ARL0: for ARL0 = 370, use $L$ of about 1.4 when
$\lambda = 0.2$, or about 1.575 when $\lambda = 0.5$; for the stricter ARL0 = 500 target, increase
$L$ by roughly 0.1 to 0.15 relative to the ARL0 = 370 value for the same $\lambda$. The chart is
robust to moderate deviations from these settings: sensitivity analysis shows that a deviation of
$\pm 0.05$ in $\lambda$, or $\pm 0.1$ in $L$, typically changes ARL0 by less than 10%. As a
balanced default, we recommend $\lambda = 0.2$ and $L = 1.4$ for ARL0 near 370, or $\lambda = 0.15$
and $L = 1.55$ for stricter control near ARL0 = 500; the illustrations above use the $(\lambda, L)
= (0.175, 1.375)$ pair from Table 4.

## Post-hoc identification: methodology and simulated performance

A chart signal indicates that the aggregate process has shifted, but it does not by itself
indicate which of the $k$ streams caused the shift. This section summarizes the statistical basis
for the package's post-hoc identification functions, the simulation evidence behind the default
choice of the Benjamini-Hochberg correction, and a worked illustration of a fundamental limit on
early-signal attribution, all drawn from Chapter 5 and Section 6.3 of [@muritala2026diss].

### Problem formulation

Consider $k$ independent streams, and let $x_{i1}, \ldots, x_{i,T_{sig}}$ be the binary indicators
for stream $i$ observed up to the signal time $T_{sig}$, where $x_{ij}=1$ if the $j$th observation
from stream $i$ exceeded the in-control median and $0$ otherwise. Under the null hypothesis that
stream $i$ is in control, $p_0 = 0.5$. Because $T_{sig}$ is determined by the aggregate chart
statistic, which is a function of all streams combined, it is independent of any single stream's
indicator sequence; conditional on $T_{sig}=t$, stream $i$'s count $S_i$ retains its
$\mathrm{Binomial}(t,p_0)$ distribution, which is what makes the p-values below valid despite the
data-dependent stopping time:

$$
S_i = \sum_{j=1}^{T_{sig}} x_{ij} \;\sim\; \mathrm{Binomial}(T_{sig},\, p_0), \qquad
p_i = P\!\big(\mathrm{Binomial}(T_{sig},p_0) \ge S_i\big) = 1 - F_{\mathrm{Bin}}(S_i - 1;\, T_{sig},\, p_0)
$$

the one-sided p-value testing for an elevated exceedance rate, which is exactly what
`calculate_pvalues()` computes using R's `pbinom()`.

### Multiple testing correction and simulation design

Testing all $k$ streams at a nominal level $\alpha$ without adjustment inflates the chance of at
least one false positive:

$$
P(\text{at least one false positive}) = 1 - (1-\alpha)^k
$$

for $k=10$ and $\alpha=0.05$, this probability reaches about 0.40. `apply_multiple_testing()`
addresses this using R's `p.adjust()` with three classical corrections, alongside the unadjusted
option:

$$
\begin{aligned}
\text{Bonferroni:} \quad & \text{reject } H_{0i} \text{ if } p_i < \alpha / k \\
\text{Holm:} \quad & \text{sort } p_{(1)} \le \cdots \le p_{(k)}; \text{ reject } H_{(1)},\ldots,H_{(j)} \text{ for the largest } j \\
& \text{such that } p_{(i)} \le \alpha/(k-i+1) \text{ for all } i \le j \\
\text{Benjamini-Hochberg:} \quad & \text{find the largest } j \text{ such that } p_{(j)} \le (j/k)\,\alpha; \text{ reject } H_{(1)},\ldots,H_{(j)}
\end{aligned}
$$

Bonferroni [@bonferroni1936] and Holm [@holm1979] both guarantee family-wise error rate (FWER) at
most $\alpha$, with Holm uniformly more powerful; Benjamini-Hochberg [@benjamini1995] instead
controls the false discovery rate (FDR), the expected proportion of flagged streams that are
actually in control, at level $\alpha$. Performance was assessed in R version 4.5.2 [@rcore2025],
parallelized with the base parallel package and doParallel, over 2,000 replications for each
combination of $\lambda$, $L$, $\delta$, and the number of truly out-of-control streams $m$
($m=0,1,2$), with the exact variance vector precomputed to $T_{max}=2{,}000$ using the
`converge_t = 500` cutoff described above. Five metrics were recorded for each method: sensitivity
(true positive rate among out-of-control streams), specificity (true negative rate among
in-control streams), FDR, FWER, and the per-stream Type I error rate. As with the parameter study
above, this 2,000-replication grid is a cited result reported from that separate simulation study,
not a computation re-run inside this document.

### Results

```{r posthocperf-table, echo=FALSE}
posthocperf_df <- data.frame(
  Method = c("Raw p-values (no correction)", "Bonferroni", "Holm", "Benjamini-Hochberg (BH)"),
  Specificity = c("about 0.96", "above 0.99", "above 0.99", "above 0.99"),
  FDR = c("0.22-0.26", "often below 0.03", "often below 0.03", "about 0.030-0.049"),
  FWER = c("above 0.20", "about 0.02 (most conservative)", "below 0.05", "below 0.05"),
  `Type I error` = c("about 0.035-0.04", "about 0.005-0.01", "about 0.005-0.01", "about 0.005-0.01"),
  `Sensitivity (rank)` = c(
    "highest, but at the cost of inflated error rates",
    "lowest among the three corrections",
    "between Bonferroni and BH",
    "highest among the three corrections"
  ),
  check.names = FALSE
)
knitr::kable(
  posthocperf_df,
  caption = "Simulated performance of the four post-hoc methods, summarized across the delta and m conditions described above. Source: Muritala (2026, Sec. 5.3).",
  booktabs = TRUE
) %>%
  kableExtra::column_spec(6, width = "1.6in") %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
```

Raw, uncorrected p-values are the most sensitive but are statistically unreliable for decision
making: their FDR runs far above the nominal 0.05 target, at roughly 0.26 when no stream is truly
out-of-control ($m=0$), 0.24 at $m=1$, and 0.22 at $m=2$, meaning that about one in four flagged
streams is actually in control, and their FWER exceeds 0.20 in all conditions studied. Bonferroni,
Holm, and BH all keep FDR and FWER near or below their nominal 0.05 targets and keep the
per-stream Type I error near zero. Among the three corrections, BH consistently shows the highest
sensitivity, followed by Holm and then Bonferroni, with the gap most visible when a single stream
is out of control ($m=1$) and the shift is moderate ($\delta$ between about 0.20 and 0.40); when
two streams are out of control ($m=2$), the three corrections perform nearly identically because
the aggregate shift is easier to detect regardless of method.

### Illustration: the minimum p-value constraint at an early signal {#sec:earlysignal}

The sensitivity figures above average over many signal times, but a signal that occurs very early
in monitoring imposes a hard constraint on attribution. If the chart signals at $T_{sig}=n$, the
raw one-sided p-value for any stream is bounded below: the smallest value attainable under
$p_0=0.5$, reached when a stream's $n$ observations are all successes, is:

$$
p_{\min}(n) \;=\; P\big(X = n \mid X \sim \mathrm{Binomial}(n, 0.5)\big) \;=\; 0.5^{\,n}
$$

Because the package's three built-in corrections, Bonferroni, Holm, and Benjamini-Hochberg, each
report an adjusted p-value that is at least as large as the corresponding raw p-value, this same
bound carries over to every correction the package offers: if $p_{\min}(n)$ already exceeds
$\alpha$, none of the four options (raw, Bonferroni, Holm, BH) can flag a stream, regardless of
the data observed. This bound is a general property of the binomial test and does not depend on
package-specific numbers, so it is verified here directly rather than only cited:

```{r pmin-check}
n <- 1:6
p_min <- 0.5^n
data.frame(n = n, p_min = p_min, below_alpha_05 = p_min < 0.05)
```

[@muritala2026diss, Sec. 6.3] works through this constraint on a simulation of $k=10$ streams over
$n=200$ time points, in which streams 1, 2, and 3 are shifted from $p_0=0.5$ to a common $p_1$
while the remaining seven streams stay in control; the chart used $\lambda=0.2$ and $L=1.4$, a
pair achieving an in-control ARL of approximately 370. Two shift sizes are reported.

Under a subtle shift ($p_1=0.58$, a shift of $\delta=0.08$), "the CSB-EWMA chart signalled a
process shift at time $t=5$. Because the signal occurred after only five time points, each stream
contributed at most five binary observations" [@muritala2026diss, Sec. 6.3.2], so the smallest
possible p-value from a binomial test on five trials, reached when all five observations are
successes, is confirmed directly above to be $p_{\min}(5) = 0.03125$. "A p-value of 0.03125 is
below the conventional significance level $\alpha=0.05$. In principle, therefore, a stream could
be flagged if it had five successes out of five. However, in this simulation the observed
successes among the shifted streams were often four or fewer, so the chart's built-in post-hoc
(which uses only the data up to the signal time) flagged no streams at $t=5$. The limitation is
practical (the observed data were not extreme enough), not theoretical" [@muritala2026diss, Sec.
6.3.2].

Under a more moderate shift ($p_1=0.65$, $\delta=0.15$), the chart signalled earlier still, at
$t=4$, where $p_{\min}(4) = 0.0625$ already exceeds $\alpha=0.05$ outright. "Therefore, no stream
could ever be flagged at the signal time, regardless of the data. As expected, the built-in
post-hoc flagged no streams" [@muritala2026diss, Sec. 6.3.2]. The two cases together show that,
under the one-sided exact binomial test used throughout this package, early-signal attribution
power is bounded by $n$ itself rather than by which of the four post-hoc options is chosen: at
$n=5$ a stream could in principle be flagged if its data were extreme enough, while at $n=4$ none
of the four options can flag any stream, however extreme the observed successes, because even the
most extreme possible outcome does not clear $\alpha=0.05$.

Applying the post-hoc procedures instead to the full 200-point history, well beyond the signal
time, gives every method enough data to work with; Table 7 reproduces the
reported result.

```{r earlysignal-table, echo=FALSE}
earlysignal_df <- data.frame(
  Method = c("Raw p-values", "Bonferroni", "Holm", "Benjamini-Hochberg"),
  `p1 = 0.58 (delta = 0.08)` = c("1, 2, 3", "2, 3", "2, 3", "1, 2, 3"),
  `p1 = 0.65 (delta = 0.15)` = c("1, 2, 3", "1, 2, 3", "1, 2, 3", "1, 2, 3"),
  check.names = FALSE
)
knitr::kable(
  earlysignal_df,
  caption = "Streams flagged by each post-hoc method when applied to the full 200-point simulated history (true out-of-control streams: 1, 2, 3). Source: Muritala (2026, Table 6.4).",
  booktabs = TRUE
)
```

"For the subtle shift ($\delta = 0.08$), Bonferroni and Holm missed one of the true
out-of-control streams (stream 1) because of their conservatism, whereas raw p-values and BH
flagged all three. For the moderate shift ($\delta = 0.15$), all methods performed equally well"
[@muritala2026diss, Sec. 6.3.2]. The second worked example above shows the same qualitative
pattern directly from the package: only two of the three shifted streams were flagged
automatically at the signal time, but a subsequent `identify_ooc()` call on the full 500-point
history flagged the third as well.

The bound $p_{\min}(n) = 0.5^n$ verified above tightens further as $n$ shrinks below 4: $p_{\min}(3)
= 0.125$, $p_{\min}(2) = 0.25$, and $p_{\min}(1) = 0.5$. Within this same one-sided binomial
framework, none of the four post-hoc options considered in this package can flag a stream on the
strength of one, two, or three post-signal observations alone: the limitation follows directly
from how little data exists immediately after an early signal, not from which correction
procedure is applied to it.

### Recommendation

Because BH keeps FDR and FWER controlled near their nominal levels while retaining the best
sensitivity of the three valid corrections, and performs at least as well as the alternatives in
the worked illustrations above, [@muritala2026diss, Sec. 5.5] recommends it as the default
post-hoc procedure, and this is why `identify_ooc()` and `csb_ewma()` both default to
`posthoc_method = "BH"`. Bonferroni or Holm remain appropriate choices when strict, guaranteed
control of the family-wise error rate is required and a corresponding loss of sensitivity is
acceptable, and the raw, unadjusted p-value option should be treated as exploratory only, since it
does not control either error rate at the nominal level. The Illustration subsection above also
illustrates a separate, practical point independent of which correction is used: when the chart
signals very early, re-running `identify_ooc()` later on more accumulated data, exactly as done in
the Illustrative use section, can recover attribution that was not yet possible at the original
signal time.

## Limitations

[@muritala2026diss, Sec. 1.7] identifies several limitations that apply directly to the package
and are worth restating for users.

- **Information loss from dichotomization.** Reducing each observation to a binary exceedance
  indicator relative to an in-control median discards magnitude information and retains only
  directional information.
- **Independence assumptions.** The method assumes independence both across streams and across
  time within a stream. In practice this is operationalized by restricting attention to weakly
  correlated streams (absolute correlation below 0.3), which can limit applicability when strong
  cross-correlation or temporal dependence is present.
- **A known, stable in-control median.** The framework requires a fixed in-control probability
  $p_0$ after dichotomization. Misspecification or drift in this baseline affects the validity of
  the resulting binomial representation and control limits.
- **Early-signal attribution power.** When the chart signals very early, few observations are
  available per stream, and the minimum attainable binomial p-value may exceed conventional
  significance thresholds, making stream-level attribution statistically infeasible regardless of
  the true shift, as derived and verified numerically in the Illustration subsection above.

Taken together, these points describe a trade-off between rapid aggregate detection and the
statistical power available for post-hoc attribution: the same short time to signal that makes the
CSB-EWMA chart attractive for early warning also limits how much can be concluded, at that same
moment, about which specific stream changed. `csb_ewma()`'s automatic post-hoc step operates only
on data available up to the signal time and will legitimately report no flagged streams in this
regime, as both the analytical bound and the worked examples above show. `identify_ooc()` can be
re-run later, once more post-signal data have accumulated, in monitoring contexts where the
process cannot be paused and continued data collection remains representative; this is not
appropriate in settings, such as manufacturing, where an out-of-control process should be
corrected immediately upon signaling.

## Summary and future work

The csbewma package makes the CSB-EWMA methodology of [@muritala2026a; @muritala2026b] and
[@muritala2026diss] directly usable in R. Its central technical contribution, an exact,
closed-form, time-varying variance for the EWMA statistic under binomial multiple-stream data,
removes the reliance on asymptotic variance approximations that limits existing binomial EWMA
charts during early-phase monitoring. The package combines this with simulation-validated
parameter guidance, an exact-inference post-hoc identification framework with Benjamini-Hochberg
as a simulation-motivated default, and ggplot2- and patchwork-based visualization, and the
Illustrative use section demonstrates the resulting workflow with real, live-evaluated package
output.

The methodology, as currently implemented, assumes independent streams and a known, stable
in-control median (see Limitations). [@muritala2026b] identify several directions for future
work: extending the framework to correlated streams, adapting it for Phase I estimation of the
in-control median and probability rather than assuming they are known, evaluating performance
under asymmetric shifts, and validating the method with further real-world case studies. From a
software perspective, natural extensions to the package include vignettes covering the workflows
illustrated above, formal unit tests under the testthat framework already declared in Suggests,
and, longer term, support for the correlated-stream and Phase I estimation extensions as the
underlying theory develops.

### Computational details

The results summarized from the underlying research were computed with R version 4.3.2
[@rcore2023] for the Monte Carlo validation and the parameter optimization, and R version 4.5.2
[@rcore2025] for the post-hoc simulation study, the latter parallelized with the base parallel
package together with doParallel; these describe the computation behind the underlying research,
not a dependency of the csbewma package itself. csbewma 1.0.1 declares R (>= 3.5) and Imports
ggplot2, stats, and patchwork in its DESCRIPTION file. All live code chunks in this article were
run with the R and package versions reported by `sessionInfo()` at the end of this document.

### Acknowledgments

We thank the members of the dissertation committee and co-authors for their guidance on the
methodology underlying this package.

```{r session-info, echo=FALSE}
sessionInfo()
```
